Some Graphical Realizations of Two-Row Specht Modules of Iwahori--Hecke Algebras of the Symmetric Group

### 2019-07-31 tex/Drafts/Draft-3.tex
addressed comments in section 4 and added comment on conjectures

  \section{Sign subrepresentations}\label{Sign section}
{\color{magenta} Smooth out the beginning of this section to work with the end of last section.}
In the previous section, we proved $M_{2n+r}^r\simeq S^{(n+r,n)}$ under certain conditions of $e$. In this proof, we used the fact that, given our conditions on $e$, we only projected down to cases with no sign subrepresentation.

In this section, we prove that there is exactly one sign subrepresentation when $e=n+r+1$, and give an explicit basis for this subrepresentation.

@@

Recall $K:= K^r_{2n+r}$, defined in section \ref{kernel}, is the direct sum of all sign subrepresentations. Here we determine an explicit basis for $K$ when $e=n+r+1$, assuming $K\not=0$. In the next section, we prove $K\not=0$.

@@

		To not be in case 1 or 2, we must have that there are no anchors between index 1 and $i$, and that there is no integer $s$ such that $1<s<i<\psi(s)-1$. It follows from the same argument that proved lemma \ref{preimage under hump} that there exists exactly one $\upsilon\in E_\psi$ such that $\upsilon(1)\not=\psi(1)$. Define $N$ to be $2n+r$ if 1 is an anchor, or $2n+r-1$ if 1 is not an anchor. Then, defining $Y_\psi$ with respect to the sub-matching $\psi(2,N)$, we have that $\sigma\in E_\psi, \sigma\not=\upsilon$ if and only if $\sigma(2,N)\in E_{\psi(2,N)}$. Note that for $E_{\psi(2,N)}$ we may apply the inductive hypothesis from case 3, so we have:
		\begin{align*}
		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_\sigma(q)\\
		=&(1+q)x_{\psi(2,N)}(q)+\sum_{\sigma\in E_{\psi(2,N)},\sigma\not=\psi(2,N)}q^{1/2}x_{\sigma}(q)\\
		=&-q^{1/2}x_{\psi(2,i-1)}x_{\psi(i+2,N)}\Q{n+r}^{i/2}
		\end{align*}

@@

		$$x_\upsilon=x_{\psi(2,i-1)}x_{\upsilon(i+2,N)}\Q{n+r}^{i/2}$$

@@

		$$-q^{1/2}x_{\psi(2,i-1)}x_{\psi(i+2,N)}\Q{n+r}^{i/2}\prn{1-1}=0$$

@@

Due to the argument present in Theorem \ref{Irreducibility Theorem}, sign subrepresentations provide a surprisingly strong characterization of some $M_{2n + r}^r$ as follows.
\begin{proposition}\label{Preimage of K}
  Fix some $M := M_{2n + r}^r$.
  Suppose there is some natural number $n'$ with $n' \leq n$ such that $K_{2n' + r}^r \neq 0$.
  Let $\pi:M_{2n + r}^r \twoheadrightarrow M_{2n' + r}^r$ be the linear map
  \[
    \pi := \begin{cases}
      (1 + T_1)(1 + T_3)\cdots(1 + T_{n-n'}) & \text{if } n < n,'\\
      \id{M} & \text{if } n = n',
    \end{cases}
  \]
  where $\id M$ is the identity map on $M$.
  Then, $\pi^{-1}(K)$ contains any proper subrepresentations of $M$.
\end{proposition}
\begin{proof}
  It suffices to show that any vector $w \in M - \pi^{-1}(K)$ is cyclic;
  then, any subrepresentation containing $w$ also contains all of $M$, implying the proposition.

@@

\subsection{Conjectures on sign subrepresentations}\label{Conjecture Section}
If $\pi^{-1}(K)$ is itself a subrepresentation of $M$, then we will have identified the radical of $M_{2n + r}^r$ via Propostion \ref{Preimage of K}.
Observing decomposition matrices for $S_{(n+r,n)}$ for small $n$, we note that $S^{(n+r,n)}$ appears to have composition series of length at-most two by \cite{Mathas-book};
the same appears to be true of $M_{2n + r}^r$, as shown in Appendix \ref{Empirics Section}.
This motivates our first conjecture:
\begin{conjecture}\label{Preimage subrepresentation}
  In Proposition \ref{Preimage of K}, $\pi^{-1}(K)$ is an irreducible subrepresentation of $M$.
\end{conjecture}

>>>>

 \section{Sign subrepresentations}\label{Sign section}
 Recall $K:=K^r_{2n+r}$, defined in section \ref{kernel}, is the direct sum of all sign subrepresentations. In the previous section, we were not able to prove $M_{2n+r}^r\simeq S^{(n+r,n)}$ for $e=n+r+1$ because we could not assume $K=0$. In particular, our proof of irreducibility failed.

 In this section, we prove that there is exactly one sign subrepresentation when $e=n+r+1$, thus $M_{2n+r}^r$ is reducible. We give an explicit basis for $K$, and explore how our approach may generalize to finding subrepresentations other $e$.

@@

 Here we determine an explicit basis for $K$ when $e=n+r+1$, assuming $K\not=0$. In the next section, we prove $K\not=0$.

@@

 		To not be in case 1 or 2, we must have that there are no anchors between index 1 and $i$, and that there is no integer $s$ such that $1<s<i<\psi(s)-1$. It follows from the same argument that proved lemma \ref{preimage under hump} that there exists exactly one $\upsilon\in E_\psi$ such that $\upsilon(1)\not=\psi(1)$. Define $N$ to be $2n+r$ if 1 is an anchor, or $2n+r-1$ if 1 is not an anchor. Then, defining $Y_\psi$ with respect to the sub-matching $\psi(2,N)$, we have that $\sigma\in E_\psi, \sigma\not=\upsilon$ if and only if $\sigma(2,N)\in E_{\psi(2,N)}$. Note that for $E_{\psi(2,N)}$ we may apply the inductive hypothesis from case 3, so we have:
 		\begin{align*}
 		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_\sigma(q)\\
 		=&(1+q)x_{\psi(2,N)}(q)+\sum_{\sigma\in E_{\psi(2,N)},\sigma\not=\psi(2,N)}q^{1/2}x_{\sigma}(q)\\
 		=&-q^{1/2}x_{\psi(2,i-1)}x_{\psi(i+2,N)}\Q_{n+r}^{i/2}
 		\end{align*}

@@

 		$$x_\upsilon=x_{\psi(2,i-1)}x_{\upsilon(i+2,N)}\Q_{n+r}^{i/2}$$

@@

 		$$-q^{1/2}x_{\psi(2,i-1)}x_{\psi(i+2,N)}\Q_{n+r}^{i/2}\prn{1-1}=0$$

@@

 Due to the argument present in Theorem \ref{Irreducibility Theorem}, sign subrepresentations provide a surprisingly strong characterization of some $M_{2n + r}^r$ as follows.
 \begin{proposition}\label{Preimage of K}
 	{\color{red!80!green}This proposition can be strengthened, right? Like we just need a sequence of $n-n'$ actions such that no two of them are closer than two positions apart, right?}
 	Fix some $M := M_{2n + r}^r$.
 	Suppose there is some natural number $n'$ with $n' \leq n$ such that $K_{2n' + r}^r \neq 0$.
 	Let $\pi:M_{2n + r}^r \twoheadrightarrow M_{2n' + r}^r$ be the linear map
 	\[
 	\pi := \begin{cases}
 	(1 + T_1)(1 + T_3)\cdots(1 + T_{n-n'}) & \text{if } n < n,'\\
 	\id{M} & \text{if } n = n',
 	\end{cases}
 	\]
 	where $\id M$ is the identity map on $M$.
 	Then, $\pi^{-1}(K)$ contains any proper subrepresentations of $M$.
 \end{proposition}
 \begin{proof}
 	It suffices to show that any vector $w \in M - \pi^{-1}(K)$ is cyclic;
 	then, any subrepresentation containing $w$ also contains all of $M$, implying the proposition.

@@

 \subsection{Conjectures on sign subrepresentations}\label{Conjecture Section}
 If $\pi^{-1}(K)$ is itself a subrepresentation of $M$, then we will have identified the radical of $M_{2n + r}^r$ via Propostion \ref{Preimage of K}.
 Observing decomposition matrices for $S_{(n+r,n)}$ for small $n$, we note that $S^{(n+r,n)}$ appears to have composition series of length at-most two by \cite{Mathas-book};
 the same appears to be true of $M_{2n + r}^r$ by Figures \ref{Composition series e=3} through \ref{Composition series e=6}.
 This motivates our first conjecture:
 \begin{conjecture}\label{Preimage subrepresentation}
 	In Proposition \ref{Preimage of K}, $\pi^{-1}(K)$ is an irreducible subrepresentation of $M$.
 \end{conjecture}
 {\color{red!80!green}Could we also say the quotient $M/\pi^{-1}(K)$ is irreducible? We believe that right?}
